Equation-of-state insensitive relations after GW170817

### 2019-02-19 paper/bin-love-after-gw170817.tex
up to Sec. 2A

@@

The structure of a NS and its tidal interactions in a binary system rely heavily on the underlying state function (or equation of state - EoS) describing the relationship between the pressure ($p$) and energy density ($\epsilon$) of nuclear matter.
Given that all currently proposed EoSs utilize certain approximations~\cite{Oertel:Review,Baym:Review}, one method to study a wide range of physically realizable EoSs is to parameterize them such that any realistic EoS can be represented with a small number of parameters.
Spectral representations~\cite{Lindblom:2010bb,Lindblom:2012zi,Lindblom:2013kra,Lindblom:2018rfr,Abbott:2018exr} parameterize EoSs by performing spectral expansions on the adiabatic index $\Gamma(p)$\footnote{Another way of parameterizing EoSs is the piecewise polytropic formulation~\cite{Read2009,Lackey:2014fwa,Carney:2018sdv}.}:
\begin{equation}
\Gamma(x) = \exp{\sum_k\gamma_k x^k},
\end{equation}
where $x \equiv \log{p/p_0}$ for a minimum pressure $p_0$.
The equation of state is then determined by an integration of the differential equation:
\begin{equation}
\frac{d \epsilon(p)}{dp}=\frac{\epsilon(p)+p}{p \Gamma(p)}.
\end{equation}
Using this formalism, any valid EoS can be approximated through the choice of 4 spectral coefficients $\gamma_k$, tabulated for several common EoSs in Table 1 of Ref.~\cite{Lindblom:2018rfr}.

In the current analysis, we consider EoSs which have been constrained by the recent binary neutron star merger event, GW170817.
Important recent work by the LIGO Collaboration~\cite{LIGO:posterior,Carney:2018sdv} sampled the EoS parameter space in order to derive a marginalized posterior on the pressure as a function of mass density, as seen in Fig. 2 of~\cite{LIGO:posterior}.
The spectral coefficients were sampled within the ranges: $\gamma_0 \in \lbrack 0.2,2 \rbrack$, $\gamma_1 \in \lbrack -1.6,1.7 \rbrack$, $\gamma_2 \in \lbrack -0.6,0.6 \rbrack$, $\gamma_3 \in \lbrack -0.02,0.02 \rbrack$, and further restricted the adiabatic index to be $\Gamma \in \lbrack 0.6,4.5 \rbrack$, ensuring the parameterization exposed a wide range of viable EoSs~\cite{Lindblom:parameters}.
Additional constraints imposed upon the generated EoSs were as follows~\cite{LIGO:posterior}: (i) causality within 10\%, and (ii) EoS priors must support NS masses up to $1.97 \text{ M}_{\odot}$, consistent with astrophysical observations~\cite{Zhao:massiveNS}.
We utilize a random set of 100 of the above posterior samples defined by GW170817, furthermore referenced as the ``constrained EoSs", as shown in Fig.~\ref{fig:eos}.
For comparison, we also include a second sample of 100 ``unconstrained" EoSs, randomly sampled in the pressure-density plane to be unrestricted by GW170817s posterior.
Following~\cite{Read2009}, the parameterized high-density core EoSs generated above are matched to the low-density crust EoS of Sly~\cite{Douchin:2001sv} at about half of the nuclear saturation density, $\rho_{\text{stitch}}=1.3 \times 10^{14} \text{ g/cm}^3$.

In addition, we investigate 10 transitional quark-hadron matter stars, which undergo first-order phase transitions at pressure $P_{\text{tr}}$, where the hadronic branch departs into a quark-matter branch at a given transitional mass~\cite{Paschalidis2018,Alford:2017qgh,1971SvA....15..347S,Zdunik:2012dj,Alford:2013aca}.
In particular, we focus on the ACS and ACB models described in Ref.~\cite{Paschalidis2018}, shown in Fig.~\ref{fig:hybridML}.
These result in two distinct types of NSs, based on their observed mass: (i) massive ($M \geq \text{ M}_{\text{tr}}$) hybrid stars which have quark-matter cores and nuclear matter crusts (we denote this as ``HS"), and (ii) low-mass ($M \leq \text{ M}_{\text{tr}}$) hadronic stars with no internal transition to quark matter (we denote this as ``NS").

>>>>

In this section we review how the EoS can be represented analytically
through a spectral decomposition, and how the observation of GW170817
constrains the space of possible EoSs. We then proceed to discuss 
how one computes the tidal deformabilities and the EoS-insensitive relations.

@@

The structure of a NS and its tidal interactions in a binary system rely heavily on the underlying state function (or equation of state - EoS) describing the relationship between the pressure ($p$) and energy density ($\epsilon$) of nuclear matter.
Given that all currently proposed EoSs utilize certain approximations~\cite{Oertel:Review,Baym:Review}, one method to study a wide range of physically realizable EoSs is to parameterize them such that any realistic EoS can be represented with a small number of parameters.
Spectral representations~\cite{Lindblom:2010bb,Lindblom:2012zi,Lindblom:2013kra,Lindblom:2018rfr,Abbott:2018exr} parameterize EoSs by performing spectral expansions on the adiabatic index $\Gamma(p)$\footnote{Another way of parameterizing EoSs is through a piecewise polytropic formulation~\cite{Read2009,Lackey:2014fwa,Carney:2018sdv}.}:
\begin{equation}
\Gamma(x) = \exp{\sum_k^{N}\gamma_k x^k},
\end{equation}
where $x \equiv \log{(p/p_0)}$ for a minimum pressure $p_0$.
The equation of state is then determined by an integration of the differential equation:
\begin{equation}
\frac{d \epsilon(p)}{dp}=\frac{\epsilon(p)+p}{p \Gamma(p)}.
\end{equation}
Using this formalism, any valid EoS can be approximated through the choice of $N$ spectral coefficients $\gamma_k$, and we here choose $N=4$, tabulated for several common EoSs in Table 1 of~\cite{Lindblom:2018rfr}.

{\ny{Zack, I rewrote what's below, but please check.}}

We here wish to consider EoS-insensitive relations using two sets of EoSs: a ``constrained set'' consistent with the observation of GW170817 and an ``unconstrained set'' that does not impose this prior. In \emph{both} cases, we model the EoS with a $C^{0}$-piecewise function that equals the low-density crust EoS of SLy~\cite{Douchin:2001sv} below half nuclear saturation density $\rho_{\text{stitch}}=1.3 \times 10^{14} \text{ g/cm}^3$~\cite{Read2009}, and equals the spectral decomposition described above outside the crust. For the latter, we restrict the spectral coefficients to the ranges $\gamma_0 \in \lbrack 0.2,2 \rbrack$, $\gamma_1 \in \lbrack -1.6,1.7 \rbrack$, $\gamma_2 \in \lbrack -0.6,0.6 \rbrack$, $\gamma_3 \in \lbrack -0.02,0.02 \rbrack$, and the adiabatic index is further restricted to $\Gamma \in \lbrack 0.6,4.5 \rbrack$~\cite{Lindblom:parameters}. Moreover, we impose the following two restrictions:  (i) causality within 10\%, i.e. that the speed of sound of the fluid be less than the speed of light to 10\%, and (ii) a high maximum mass, i.e.~that the resulting EoS supports NSs with masses at least as high as $1.97 \text{ M}_{\odot}$, consistent with astrophysical observations~\cite{Zhao:massiveNS}.

The unconstrained set is then defined by drawing random samples in the spectral coefficients (within their allowed prior ranges), and then eliminating any EoS that either leads to a polytropic index $\Gamma$ outside the allowed range, or breaks the causality restriction, or breaks the maximum mass restriction. The constrained set is defined in a similar way, with the addition of the restriction that the resulting EoS be inside the 90\% marginalized posterior distribution on the pressure-mass density plane derived from the GW170817 observation~\cite{LIGO:posterior,Carney:2018sdv}  (see Fig.~2 in~\cite{LIGO:posterior}). In both cases, each set consists of 100 members, with a subset of these shown in Fig.~\ref{fig:eos}.

%In the current analysis, we consider EoSs which have been constrained by the recent binary neutron star merger event, GW170817.
%Important recent work by the LIGO Collaboration~\cite{LIGO:posterior,Carney:2018sdv} sampled the EoS parameter space in order to derive a marginalized posterior on the pressure as a function of mass density, as seen in Fig. 2 of~\cite{LIGO:posterior}.
%The spectral coefficients were sampled within the ranges: $\gamma_0 \in \lbrack 0.2,2 \rbrack$, $\gamma_1 \in \lbrack -1.6,1.7 \rbrack$, $\gamma_2 \in \lbrack -0.6,0.6 \rbrack$, $\gamma_3 \in \lbrack -0.02,0.02 \rbrack$, and further restricted the adiabatic index to be $\Gamma \in \lbrack 0.6,4.5 \rbrack$, ensuring the parameterization exposed a wide range of viable EoSs~\cite{Lindblom:parameters}.
%Additional constraints imposed upon the generated EoSs were as follows~\cite{LIGO:posterior}: (i) causality within 10\%, and (ii) EoS priors must support NS masses up to $1.97 \text{ M}_{\odot}$, consistent with astrophysical observations~\cite{Zhao:massiveNS}.
%We utilize a random set of 100 of the above posterior samples defined by GW170817, furthermore referenced as the ``constrained EoSs", as shown in Fig.~\ref{fig:eos}.
%For comparison, we also include a second sample of 100 ``unconstrained" EoSs, randomly sampled in the pressure-density plane to be unrestricted by GW170817s posterior.
%Following~\cite{Read2009}, the parameterized high-density core EoSs generated above are matched to the low-density crust EoS of Sly~\cite{Douchin:2001sv} at about half of the nuclear saturation density, $\rho_{\text{stitch}}=1.3 \times 10^{14} \text{ g/cm}^3$.

{\ny{Zack, I think the description here was wrong because the hybrid stars do not just have hadronic matter in the crust, but also in the outer core, no? Please check.}}

In addition, we investigate 10 transitional quark-hadron matter stars, which undergo strong first-order phase transitions at a pressure $P_{\text{tr}}$, leading to the hadronic branch departing into a quark-matter branch at a given transitional mass~\cite{Paschalidis2018,Alford:2017qgh,1971SvA....15..347S,Zdunik:2012dj,Alford:2013aca}. In particular, we focus on the ACS and ACB models described in~\cite{Paschalidis2018}, and also shown in Fig.~\ref{fig:hybridML}. These result in two distinct types of NSs, based on their mass: (i) massive ($M \geq \text{ M}_{\text{tr}}$) hybrid stars which have quark-matter inner cores and nuclear matter elsewhere (henceforth, we denote such stars as hybrid stars (HSs)), and (ii) low-mass ($M \leq \text{ M}_{\text{tr}}$) hadronic stars with no internal transition to quark matter (henceforth, we denote these stars as simply NSs).
